---
title: "H3K27ac_HGSOC_consensus_upsetr_analysis"
author: "Alberto Luiz Pascasio Reyes"
date: "10/25/2018"
output: html_document
---

load libraries and input files
```{r}

library(GenomicRanges)
library(GenomicFeatures)
library(HelloRanges)
library(IRanges)
library(S4Vectors)
library(readr)
library(dplyr)
library(tidyr)
library(tidyverse)
library(ggplot2)
library(UpSetR)
library(rtracklayer)
library(optparse)
library(pheatmap)

hgsoc_k27ac_hg19 <- read_table2("hg19_H3K27ac_biofeature_beds/HGSOC/allpeaks.anno.tsv") ### file made from simon's bash script

```

form peak subsets: unique, consensus, strict
```{r}

### make union of all unique peaks
union_unique_hgsoc_k27ac_hg19 <- hgsoc_k27ac_hg19[which(rowSums(hgsoc_k27ac_hg19[,4:7]) < 1.5),-8] ##### NOT merged

### pull out cell line specific unique peaks to make data frames
hey_unique_hgsoc_k27ac_hg19 <- filter(union_unique_hgsoc_k27ac_hg19, B08_Hey_H3K27Ac_hg37 == 1)
uwb_unique_hgsoc_k27ac_hg19 <- filter(union_unique_hgsoc_k27ac_hg19, B09_UWB1.289 == 1)
caov3_unique_hgsoc_k27ac_hg19 <- filter(union_unique_hgsoc_k27ac_hg19, B10_CaOV3 == 1)
kur_unique_hgsoc_k27ac_hg19 <- filter(union_unique_hgsoc_k27ac_hg19, B11_Kuramochi_H3K27Ac_hg37 == 1) ##### *** CONFIRM "NEW PEAKS" ARE 2017 PANEL LIFTED OVER FROM HG38 TO HG19 --> ALL BIOFEATURE FILES STILL HG19??? ***

### convert into granges and ignore metadata cosl (bc that information will be in name of peak set)
gr_hey_unique_hgsoc_k27ac_hg19 <- makeGRangesFromDataFrame(hey_unique_hgsoc_k27ac_hg19, keep.extra.columns = FALSE)
gr_uwb_unique_hgsoc_k27ac_hg19 <- makeGRangesFromDataFrame(uwb_unique_hgsoc_k27ac_hg19, keep.extra.columns = FALSE)
gr_caov3_unique_hgsoc_k27ac_hg19 <- makeGRangesFromDataFrame(caov3_unique_hgsoc_k27ac_hg19, keep.extra.columns = FALSE)
gr_kur_unique_hgsoc_k27ac_hg19 <- makeGRangesFromDataFrame(kur_unique_hgsoc_k27ac_hg19, keep.extra.columns = FALSE)
# gr_hey_unique_hgsoc_k27ac_hg19 <- GRanges(seqnames = hey_unique_hgsoc_k27ac_hg19$chr,
#                                           ranges = IRanges(start = hey_unique_hgsoc_k27ac_hg19$start, end = hey_unique_hgsoc_k27ac_hg19$end -1),
#                                           data = NULL)


# ### merge individual GR obj's
# gr_hey_unique_hgsoc_k27ac_hg19
# bedtools_merge("-i hey.bed")
# gr_hey_unique_hgsoc_k27ac_hg19 <- GenomicRanges::reduce(gr_hey_unique_hgsoc_k27ac_hg19, ignore.strand = TRUE, with.revmap = TRUE)
# gr_hey_unique_hgsoc_k27ac_hg19 ####### --> do NOT need to do merge step for unique (duh)???
# 
# gr_uwb_unique_hgsoc_k27ac_hg19
# gr_uwb_unique_hgsoc_k27ac_hg19 <- GenomicRanges::reduce(gr_uwb_unique_hgsoc_k27ac_hg19, ignore.strand = TRUE)
# gr_uwb_unique_hgsoc_k27ac_hg19 ####### --> do NOT need to do merge step for unique (duh)???



### make consensus peak set df
consensus_hgsoc_k27ac_hg19 <- hgsoc_k27ac_hg19[which(rowSums(hgsoc_k27ac_hg19[,4:7]) >= 1.5), -8] ##### NOT merged
### ********* THIS IS WRONG BC CAN BE 1 + 0.3 _ 0.2 + 0 *********

### convert into granges and ignore metadata cols
gr_consensus_hgsoc_k27ac_hg19 <- makeGRangesFromDataFrame(consensus_hgsoc_k27ac_hg19, keep.extra.columns = FALSE)

### merge consensus GR obj (don't double count regions/"peaks")
gr_consensus_hgsoc_k27ac_hg19 ### 136361 peaks
gr_consensus_hgsoc_k27ac_hg19 <- GenomicRanges::reduce(gr_consensus_hgsoc_k27ac_hg19, ignore.strand = FALSE)
gr_consensus_hgsoc_k27ac_hg19 ### 39014 peaks --> douesn't double count overlapping regions



### make strict subset of peaks (peaks that ONLY APPEAR IN ALL SAMPLES) df
# strict_hgsoc_k27ac_hg19 <- hgsoc_k27ac_hg19[which(rowSums(hgsoc_k27ac_hg19[4:7]) > 3), -8]



hgsoc_k27ac_hg19 %>% select(B08_Hey_H3K27Ac_hg37, B09_UWB1.289, B10_CaOV3, B11_Kuramochi_H3K27Ac_hg37) %>% 
  filter(B08_Hey_H3K27Ac_hg37 >= 0.5 & B09_UWB1.289 >= 0.5 & B10_CaOV3 >= 0.5 & B11_Kuramochi_H3K27Ac_hg37 >= 0.5)

strict_hgsoc_k27ac_hg19 <- hgsoc_k27ac_hg19 %>% select(chr, start, end, B08_Hey_H3K27Ac_hg37, B09_UWB1.289, B10_CaOV3, B11_Kuramochi_H3K27Ac_hg37) %>% 
                              filter(B08_Hey_H3K27Ac_hg37 >= 0.5 & B09_UWB1.289 >= 0.5 & B10_CaOV3 >= 0.5 & B11_Kuramochi_H3K27Ac_hg37 >= 0.5)

strict_hgsoc_k27ac_hg19_gr <- makeGRangesFromDataFrame(strict_hgsoc_k27ac_hg19, keep.extra.columns = TRUE)

strict_gr_r <- GenomicRanges::reduce(strict_hgsoc_k27ac_hg19_gr, with.revmap = TRUE)


```

write above peak subsets to files
```{r}

### write consensus
View(as.data.frame(gr_consensus_hgsoc_k27ac_hg19))
write.table(as.data.frame(gr_consensus_hgsoc_k27ac_hg19),
            file = "./peak_subsets/hg19_H3K27ac/HGSOC/hgsoc_k27ac_hg19_consensus.bed",
            row.names = FALSE,
            col.names = FALSE,
            quote = FALSE,
            sep = "\t")

### write hey unique
write.table(as.data.frame(gr_hey_unique_hgsoc_k27ac_hg19),
            file = "./peak_subsets/hg19_H3K27ac/HGSOC/unique_hey_hgsoc_k27ac_hg19.bed",
            row.names = FALSE,
            col.names = FALSE,
            quote = FALSE,
            sep = "\t")

### write uwb unique
write.table(as.data.frame(gr_uwb_unique_hgsoc_k27ac_hg19),
            file = "./peak_subsets/hg19_H3K27ac/HGSOC/unique_uwb_hgsoc_k27ac_hg19.bed",
            row.names = FALSE,
            col.names = FALSE,
            quote = FALSE,
            sep = "\t")

### write caov3 unique
write.table(as.data.frame(gr_caov3_unique_hgsoc_k27ac_hg19),
            file = "./peak_subsets/hg19_H3K27ac/HGSOC/unique_caov3_hgsoc_k27ac_hg19.bed",
            row.names = FALSE,
            col.names = FALSE,
            quote = FALSE,
            sep = "\t")

### write kuramochi unique
write.table(as.data.frame(gr_kur_unique_hgsoc_k27ac_hg19),
            file = "./peak_subsets/hg19_H3K27ac/HGSOC/unique_kur_hgsoc_k27ac_hg19.bed",
            row.names = FALSE,
            col.names = FALSE,
            quote = FALSE,
            sep = "\t")



```

do consensus/intersection/upsetr analysis
NOTE: consider how to NOT do bedtools merge step to NOT deflate # of shared peaks
```{r}

# option_list <- list(
#   make_option(c("-f", "--file"), type = "character", default = NULL,
#               help = "dataset file name", metavar = "character"),
#   make_option(c("-o", "--out"), type = "character",
#               default = "consensus_out",
#               help = "output file name [default= %default]",
#               metavar = "character")
# )
# 
# opt_parser <- OptionParser(option_list = option_list);
# opt <- parse_args(opt_parser);
# 
# if (is.null(opt$file)){
#   print_help(opt_parser)
#   stop("At least one argument must be supplied (input file).n", call.=FALSE)
# }
# 
# if (!file.exists(opt$file)) {
#   stop("File ", opt$file, "does not exist")
# } ### not working yet

x <- hgsoc_k27ac_hg19[,-8] ###change this obj name to whatever file being analyzed
View(x)

y <- x; y[, c(4:ncol(x))] <- x[, c(4:ncol(x))] > 0.5
View(y)

y <- GRanges(seqnames = y$chr,
             ranges = IRanges(start = y$start, end = y$end - 1),
             data = y[, 4:ncol(y)])
y

yr <- GenomicRanges::reduce(y, with.revmap = TRUE) ### try excluding this step to NOT merge and undercount shared peaks

sample_overlap <- t(bind_cols(lapply(
  relist(mcols(y)[unlist(yr$revmap),], yr$revmap),
  function(x) {
    tibble::as_tibble(colSums(as.matrix(x)) >= 1) ### set threshold of overlap (percentage) here.....why is this mat bigger than x???...how to interp this chunk
  })))

mcols(yr) <- sample_overlap
colnames(sample_overlap) <- gsub("data.", "", colnames(mcols(y)))
yr$sample_count <- rowSums(as.matrix(mcols(yr)))

# file_ranges <- yr
# save(file_ranges, sample_overlap, file = paste0(opt$out, "_rdata.rda"))
# 
# pdf(paste0(opt$out, "_upset.pdf")) ### not yet working

upset(as.data.frame(sample_overlap + 0), ### simon's method
      order.by = "freq",
      text.scale = c(2, 2, 2, 1.5, 2.25, 2.25),
      point.size = 5) ### use this method bc of reduce step (ie bedtools merge) --> doesn't double count peaks, instead uses our defn of inclusive boundaries
######### looks very different than first correction attempt (but this one is correct bc last time used WRONG input files)

# dev.off()
# 
# for (peak_count in c(1:ncol(mcols(y)))) {
#   rtracklayer::export.bed(yr[yr$sample_count == peak_count, ],
#                           paste0(opt$out, "_", peak_count, "_samples.bed"))
# } ### not working yet

rm(z)
z <- (hgsoc_k27ac_hg19[,4:7] >= 0.5) + 0
View(z)
upset(as.data.frame(z), order.by = "freq")

```



above method for calling consensus peaks (rowSums >= 1.5) OVERCOUNTS bc possible 1 + 0.3 + 0.2 which should NOT be called consensus
```{r}

### write fct to check that AT LEAST ONE other col is >=0.5
### input "consensus" matrix
### check each col for 1, once found then note TRUE (should be true for all rows)
### check each col for 0.5, once found then note TRUE (if true then consensus if NOT then exclude)

check_consensus <- function(filtmat, minov) ### filt mat: filtered matrix (filter by row Sum) 
{
  # for(i in 1:dim(filtmat)[1])
  # {
  #   for(j in 4:dim(filtmat)[2])
  #   {
  #     
  #   }
  # }
  
  filtmatbin <- cbind(filtmat[,1:3], (filtmat[,4:7] >= minov) + 0) ### filter matrix binary
  
  conmat <- filtmatbin[which(rowSums(filtmatbin[,4:7]) >= 2),]
  
  conmat
  
}

zz <- check_consensus(consensus_hgsoc_k27ac_hg19, 0.5) ### this now matches row number of NON-MERGED input upsetr plot

nrow(zz[which(rowSums(zz[,4:7]) ==4),])

```


try cluster analysis
```{r}

pheatmap(as.matrix(x[1:1000,4:7])) ### clustering "works" but not meaningful w/ just presence of peak (esp. bc all peaks used)???

```

"cluster" (using peak overlap/presence) using consensus set then unique sets?
```{r}

x <- hgsoc_k27ac_hg19[which(rowSums(hgsoc_k27ac_hg19[,4:7]) >= 1.5), -8] ##### NOT merged.....but can't if doing the following steps *** redefining x ***
pheatmap(as.matrix(x[runif(1000, 1, 136361) ,4:7])) ###### NO MEANING doing hgsoc k27ac consensus by itself bc no compare to other groups --> try unique set

x <- filter(union_unique_hgsoc_k27ac_hg19, B08_Hey_H3K27Ac_hg37 == 1)
pheatmap(as.matrix(x[runif(1000, 1, 18990) ,4:7])) ### behaves as expected

x <- hgsoc_k27ac_hg19[which(rowSums(hgsoc_k27ac_hg19[,4:7]) < 1.5),-8]
pheatmap(as.matrix(x[runif(1000, 1, 106698) ,4:7]))

##### need to do similar to all consensus w/ each other #####

```



make TSS file.....attempt
```{r}

t1 <- GenomicFeatures::makeTranscriptDbFromUCSC("hg38", "refGene") ##### getting error

```


